---
title: "Analysis"
author: "Lavonnia Newman, Joseph Caguioa, Jeff Washburn"
date: "12/7/2018"
output: 
  html_document:
    theme: default
    highlight: tango
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false

---

```{r setup, include=FALSE, message=F}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("psych")
library(knitr)
library(ggplot2)
library(dplyr)
library(plyr)
library(psych)

```

# Placeholder. RMarkdown file should contain:
* Formal introduction to client.
* Code explanations as sentences before or after chunks.
* Explicit answers to following 7 questions, given in sentences outside code chunks.
* Project conclusion summarizing findings.

# Data Cleansing

``` {r clean}

# Create new working copies.
beers_raw <- read.csv("./Analysis/Beers.csv", header=TRUE, sep=",")
breweries_raw <- read.csv("./Analysis/Breweries.csv", header=TRUE, sep=",")

# Create new working copies.
beers_clean <- data.frame(beers_raw)
breweries_clean <- data.frame(breweries_raw)

# Change column name of brewery ID to match for merging.
names(beers_clean)[names(beers_clean) == "Brewery_id"] <- "Brewery_ID"
names(beers_clean)[names(beers_clean) == "Name"] <- "Beer_Name"
names(breweries_clean)[names(breweries_clean) == "Brew_ID"] <- "Brewery_ID"
names(breweries_clean)[names(breweries_clean) == "Name"] <- "Brewery_Name"

# Change data types of Names to character. Style, City, and State stay factors.
beers_clean$Beer_Name <- as.character(beers_clean$Beer_Name)
breweries_clean$Brewery_Name <- as.character(breweries_clean$Brewery_Name)

```

# The questions:
## 1. How many breweries are present in each state? 
(Alongside a nicely formatted table, consider a choropleth map if possible.)

``` {r numbreweries}

# How many breweries are there per state? Uses count() from plyr.
breweries_count_state <- count(breweries_clean, "State")
names(breweries_count_state)[names(breweries_count_state) == "freq"] <- c("Num_Breweries")

# Bar graph visualization.
bps_bars <- breweries_count_state %>% ggplot(aes(x = reorder(State, Num_Breweries), y = Num_Breweries)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("Number of Breweries Per State") +
  xlab("State") +
  ylab("Number of Breweries") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks = element_blank()
        )
bps_bars

```

## 2. Merge beer data with breweries data. 
Print first 6 observations and last 6 observations to check merged file.

```{r merge}

merged_beers_raw <- merge(beers_clean, breweries_clean, by = c("Brewery_ID", "Brewery_ID"), all = TRUE)
merged_beers_raw <- merged_beers_raw[c("Beer_Name", "Beer_ID", "ABV", 
                               "IBU", "Style", "Ounces",
                               "Brewery_Name", "Brewery_ID", "City", "State")]

kable(head(merged_beers_raw, 6), format="markdown")
kable(tail(merged_beers_raw, 6), format="markdown")

```

## 3. Report number of NAs in each column.

```{r numnas}

# Report the number of NAs in each column.
num_nas <- colSums(is.na(merged_beers_raw))
num_nas
```

## 4. Compute median alcohol content and international bitterness unit for each state. 
Plot a bar chart to compare.

```{r medians_state}
#Calculating the median and ignoring NA
abv_state_median<-aggregate(merged_beers_raw$ABV, list(merged_beers_raw$State), median, na.rm=TRUE)
ibu_state_median<-aggregate(merged_beers_raw$IBU, list(merged_beers_raw$State), median, na.rm=TRUE)

names(abv_state_median)<-c("State", "Median ABV")
names(ibu_state_median)<-c("State", "Median IBU")
```
### 4.1 ABV Median
``` {r}
#Output the results
kable(abv_state_median, format="markdown")
```
### 4.2 IBU Median
``` {r}
kable(ibu_state_median, format="markdown")
```

## 5. Which state has maximum ABV beer? Most bitter beer in IBU? 
(Consider bar chart again, with top [5/10] in each category. However, highlight the number 1.)
```{r "abv-ibu echo=TRUE"}
abv_max <- max(merged_beers_raw$ABV, na.rm=TRUE)
ibu_max <- max(merged_beers_raw$IBU, na.rm=TRUE)
abv_max
ibu_max

#Convert State from Factor to Character
merged_beers_raw$State <- as.character(merged_beers_raw$State)

#Determine the highest ABV in each State (return by State)
abv.agg <- aggregate(ABV ~ State, merged_beers_raw, max)

#Now Sort the states by ABV descending
sorted_abv_agg<-abv.agg %>% arrange(desc(ABV))
first_ten_abv<-head(sorted_abv_agg, 10)
kable( head(first_ten_abv, 10), format="markdown")


#Determine the highest IBU in each State (return by State)
ibu.agg <- aggregate(IBU ~ State, merged_beers_raw, max)

#Now Sort the states by IBU descending
sorted_ibu_agg<-ibu.agg %>% arrange(desc(IBU))
kable( head(sorted_ibu_agg, 10), format="markdown")


#Now let's graph ABV (top 10)
abv_beers <- first_ten_abv %>% ggplot(aes(x= reorder(State, desc(ABV)), y = ABV)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("States by ABV") +
  xlab("State") +
  ylab("ABV") +
  theme(plot.title = element_text(hjust = 0.5), axis.ticks = element_blank()) + 
  theme(axis.text.x = element_text(angle = -90))
abv_beers
```

## 6. Summary statistics for ABV.
```{r "abv-summary-stats echo=TRUE"}
summary(merged_beers_raw$ABV)
```

## 7. Is there a relationship between IBU and ABV? 
Draw a scatterplot. (Use linear regression to quantify the relationship, provide R^2.)
```{r "IBU-to-ABV echo=TRUE"}
abv_ibu_scatter <-
 merged_beers_raw %>% ggplot(aes(x = ABV, y = IBU)) +
 geom_point(na.rm = TRUE, alpha = 0.5, col="#F19800") +
 xlab("Alcohol By Volume (ABV)") +
 ylab("International Bitterness Units (IBU)") +
 ggtitle("Relationship Between Beer’s\nBitterness and Alcoholic Content") +
 # Test with both method="lm" and ="loess" for linear and smooth regressions.
 geom_smooth(method = "lm", se = TRUE, level = 0.95,
             aes(color = "CC720C", fill = "#F28E1C")
             ) +
 theme(plot.title = element_text(hjust = 0.5),
       axis.ticks = element_blank()
       )
abv_ibu_scatter
```
Results - There appears to be a positive correlation between ABV and IBU.  The higher the alcoholic content (ABV) the higher the bitterness (IBU).  

# Storyline
## Notes
* Concept
  + Dr Cheun is an established bar owner, looking to open a new bar(s) across the U.S.  We are an analysist firm helping Dr. Cheun open bars.  
  + She is looking for:
    + Open a new bar, that has easy access to most breweries for the type of beers to target certain customers
    + Types of beer
    + Bar location (Dr. Cheun is based out of Nevada, we think you should not open up a new bar here)
    + Range of beer (which beers to have on tap)
    + Need to map current locations of her existing bars (to reconfirm current selection of beers against consumer base)
      + We analyzed our beers from ABV and IBU content and matched up the best selection based on your store location.
      + Beer xzy would do good here vs, beer def will do better in your TX bar because they like to get drunk quick
    + Demographics of this state is mostly xyz, and found this beer is a good fit for this abc beer
* Company Name
  + SM-Brew (new name in the works)
* Call out that Dr Cheun is the client (gave use the data to be analyzed) and we are the vendor/sales/ folks that took data, cleaned, merged it, made sense of it
  + 2 data files provided, had to merge them. 
  + in IBU and ABV, there was x number of these when we merged them by beers across states, there was x number of NAs for number of beers.  
  + We cleaned some things
* Content that must be worked into storyline
  + #1 Breweries per state
  + #4 Median ABV and IBU per state
  + #5 State w/ max ABV beer and most bitter 
  + #6 Summary statistic for ABV
  + #7 Relationship between ABV and IBU
* Cast
  + Dr. Cheun investor 
  
## Style Analysis

```{r style_analysis}

# General counts of beer styles. Obtain top 20 by brewing frequency.
beers_count_style <- count(merged_beers_raw, "Style")
names(beers_count_style)[names(beers_count_style) == "freq"] <- c("Style_Count")
beers_count_style_sorted <- beers_count_style %>% arrange(desc(Style_Count))
beers_count_style_sorted_top20freq <- beers_count_style_sorted[1:20, ]

# Total number of beer styles is 100.
num_styles <- n_distinct(beers_count_style$Style)
cat("The original data contains ", num_styles, " unique beer styles.")
# num_styles_check <- n_distinct(beers_raw$Style)
# cat("The raw version of beers contains ", num_styles_check, " unique beer styles. Does this match the other statement?")

kable(beers_count_style_sorted_top20freq, format="markdown")

# Bar graph visualization.
style_bars <- beers_count_style_sorted_top20freq %>% ggplot(aes(x = reorder(Style, Style_Count), y = Style_Count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("Top 20 Styles by Frequency") +
  xlab("Style") +
  ylab("Beers Per Style") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks = element_blank()
        )
style_bars

```

``` {r brewery_analysis}

# Count instances of Brewery_Name to figure out how many beers each brewery makes.
beers_in_brewery <- count(merged_beers_raw, "Brewery_Name")
names(beers_in_brewery)[names(beers_in_brewery) == "freq"] <- c("Beer_Count")
beers_in_brewery <- beers_in_brewery %>% arrange(desc(Beer_Count))
beers_in_brewery_top20freq <- beers_in_brewery[1:20, ]
kable(beers_in_brewery_top20freq, format="markdown")

```

## Intro
## 3 main points
### Point 1
### Point 2
### Point 3

## Conclusion

