---
title: "Analysis"
author: "Joseph Caguioa"
date: "12/7/2018"
output: 
  html_document:
    theme: simplex
    highlight: zenburn
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false

---

```{r setup, include=FALSE, message=F}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(ggplot2)
library(dplyr)
library(plyr)
library(psych)

```

# Placeholder. RMarkdown file should contain:
* Formal introduction to client.
* Code explanations as sentences before or after chunks.
* Explicit answers to following 7 questions, given in sentences outside code chunks.
* Project conclusion summarizing findings.

# Data Cleansing

``` {r clean}

# Create new working copies.
beers_raw <- read.csv("./Analysis/Beers.csv", header=TRUE, sep=",")
breweries_raw <- read.csv("./Analysis/Breweries.csv", header=TRUE, sep=",")

# Create new working copies.
beers_clean <- data.frame(beers_raw)
breweries_clean <- data.frame(breweries_raw)

# Change column name of brewery ID to match for merging.
names(beers_clean)[names(beers_clean) == "Brewery_id"] <- "Brewery_ID"
names(beers_clean)[names(beers_clean) == "Name"] <- "Beer_Name"
names(breweries_clean)[names(breweries_clean) == "Brew_ID"] <- "Brewery_ID"
names(breweries_clean)[names(breweries_clean) == "Name"] <- "Brewery_Name"

# Change data types of Names to character. Style, City, and State stay factors.
beers_clean$Beer_Name <- as.character(beers_clean$Beer_Name)
breweries_clean$Brewery_Name <- as.character(breweries_clean$Brewery_Name)

```

# The questions:
## 1. How many breweries are present in each state? 
(Alongside a nicely formatted table, consider a choropleth map if possible.)

``` {r numbreweries}

# How many breweries are there per state? Uses count() from plyr.
breweries_count_state <- count(breweries_clean, "State")
names(breweries_count_state)[names(breweries_count_state) == "freq"] <- c("Num_Breweries")

# Bar graph visualization.
bps_bars <- breweries_count_state %>% ggplot(aes(x = reorder(State, Num_Breweries), y = Num_Breweries)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("Number of Breweries Per State") +
  xlab("State") +
  ylab("Number of Breweries") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks = element_blank()
        )
bps_bars

```

## 2. Merge beer data with breweries data. 
Print first 6 observations and last 6 observations to check merged file.

```{r merge}

merged_beers <- merge(beers_clean, breweries_clean, by = c("Brewery_ID", "Brewery_ID"), all = TRUE)
merged_beers <- merged_beers[c("Beer_Name", "Beer_ID", "ABV", 
                               "IBU", "Style", "Ounces",
                               "Brewery_Name", "Brewery_ID", "City", "State")]

kable(head(merged_beers, 6), format="markdown")
kable(tail(merged_beers, 6), format="markdown")

```

## 3. Report number of NAs in each column.

```{r numnas}

# Report the number of NAs in each column.
num_nas <- colSums(is.na(merged_beers))
num_nas

```

## 4. Compute median alcohol content and international bitterness unit for each state. 
Plot a bar chart to compare.

```{r medians_state}



```

## 5. Which state has maximum ABV beer? Most bitter beer in IBU? 
(Consider bar chart again, with top [5/10] in each category. However, highlight the number 1.)

## 6. Summary statistics for ABV.

## 7. Is there a relationship between IBU and ABV? 
Draw a scatterplot. (Use linear regression to quantify the relationship, provide R^2.)